Title here

### 2016-12-16 portfolio/text/ms.Rmd
Edits incorporating Ole's edits

# Abstract
Ole would drop the confidence bounds for the abstract.  I added a bit but this seems like a long abstract for PNAS...

@@

Our results suggest that thebenefit of species diversity for revenue variability is mostly an across-permit
effect, not a benefit that can be easily harnessed by an individual permit
holder. Effective methods of diversification --- purchasing additional permits
or switching permits --- are costly and often limited by regulations such as
limited entry programs and individual transferable quotas (ITQs). Take home ...

@@

Commercial fishers present an interesting example of diversification, given the
extreme risk that some individuals experience, particularly relative to
terrestrial agriculture [@sethi2012b]. The majority of risk insurance programs
available to terrestrial farmers (e.g. crop insurance programs) are not
obtainable by commercial fishers [@kasperski2013], possibly because both fish
and fishers are spatially mobile 
[Ole is pretty sure it is because they don't have sufficient political clout, suggest deleting the clause after the comma in previous sentence]. 
However, fishers maintain the ability to
lower risk by diversifying their fishing portfolio. They can, for example, target as many species as permitted by
a license, own multiple licenses, or diversify spatially and fish in multiple
regions. The effectiveness of these diversification strategies for increasing
revenue and reducing revenue variability remains uncertain. For example, theory
and empirical studies suggest there can be substantial benefits to
specialization for both farmers and fishers [@abdulai2000; @arnason2005].
Furthermore, the majority of research examining the effects of diversification
on revenue do so at aggregate levels that combine individuals (e.g. households,
counties, or communities; [@kasperski2013; @sethi2014]). Thus, key
uncertainties remain about the fundamental relationship between diversity,
fishing revenue, and fishing revenue variability at the level of the individual.

@@

We develop a variance-function regression model [@davidian1987; @hall1989],
to simultaneously models changes in the mean and variance of individual
revenue as a function of covariates. [OLE: I think we leave this sentence for the methods somewhere- Whereas traditional hierarchical linear models assume homoscedasticity --- that the residual variance does not vary
systematically --- the variance-function regression model allows the residual
variance to be heteroscedastic and vary as a function of predictors.] 
We simultaneously estimate the effect of species diversity (the
inverse of Simpson's diversity index [@simpson1949; @jost2006]) on individual
revenue and revenue variability, and estimate effects both within and across
fishing strategies while controling for fishing effort and latent time effects 
(e.g. changing environmental conditions). To our knowledge, this is the first application of this type of
model in any natural resource field.

@@

However, adjusting (increasing or decreasing) an
individual's fished species diversity within the constraints of a permit set
tends to be associated with greater expected variability or uncertainty in
year-to-year revenue (Fig. \ref{fig:within} strategies above 0; Figs.
S\ref{fig:fe}, \ref{fig:re}, \ref{fig:mirror}). This effect was strongest for
individuals with a halibut permit or halibut and sablefish permits. Individuals
with these strategies experienced ~2% increase in variability for a 1% change in
the year-to-year ratio of species diversity (Fig. \ref{fig:within}). Interestingly, becoming more generalized in species diversity also leads to increased revenue variability (FIG SUPP), suggesting that the act of changing strategies is a contributor to increased variability of revenue.

Levels of fished species diversity above one for these particular permits are primarily
due to bycatch.

[OLE: I would just say - see supplement for discussion of consequences of generalizing and not include this section in the main text]
The converse is also usually true --- generalizing
tends to be associated with lower expected revenue after accounting for fishing
effort (i.e. Fig. S\ref{fig:mirror}A and S\ref{fig:mirror}B are near mirror
images left to right). This benefit of specializing is most apparent after
accounting for fishing effort. Given that fishing for more species tends to be
associated with increased fishing effort (fishing more days), specializing tends
to be ...TODO (Fig. S\ref{fig:noeffort}).

Alaskan halibut and sablefish fisheries were converted to individual fishery
quota (IFQ) management in 1994, which potentially confounds our estimates of the effect of
diversity and a major regulatory change. We refit our models allowing all strategy-level effects
to vary separately before and after the introduction of IFQs for strategies involving halibut
and sablefish permits. The introduction of IFQs reduced variability by ~25% (XX--XX 95% CI)
for individuals with only halibut or sablefish permits (Fig. S\ref{fig:ifq}A). 
Furthermore, for most strategies, the association changing species diversity with increased
revenue variability became stronger after the introduction of IFQs (Fig.
S\ref{fig:ifq}B). This was especially true for individuals fishing only halibut
or sablefish and less so for individuals combining halibut or sablefish with
a salmon permit (Fig. S\ref{fig:ifq}B).

# Discussion
[Ole thinks we should start bigger here to reflect the big main themes in the intro]

For individuals deriving livelihoods from natural resources, theory suggests that adopting strategies that utilize
a diversity of assets [species?] provide substantial benefits over strategies with lower diversity [REFS].
However, the ability to effectively diversify in practice is limited by regulation and market forces.
Here we document the effect of changing species diversity on revenue and revenue variability across 
three scale. Each scale reflects individual level choices that produce varied levels of species diversity.  
We find that individuals fishing permit sets that allow high levels
of species diversity tend to have lower revenue variability. Adding an
additional fisheries permit usually increases revenue and decreases revenue
variability. However, within the constraints of a permit and accounting for
effort, specializing (reducing species diversity) tends to increase revenue and
specializing or generalizing tends to be associated with greater revenue
variability, with considerable differences across fisheries. Thus, the most
effective mechanism by which species diversity can lower revenue variability for
individual fishers --- switching or adding fishing permits --- is often costly,
limited by regulations, and consequently unavailable to many individuals in practice.

Our results shows additional support for the idea that individual fishers can reduce 
their income variability by diversifying their catch portfolios. However, most individuals
don’t appear to change their catch diversity significantly from year to year, and 
most strategies adopted by Alaskan fishermen lie in a high-risk, high-reward quadrant
in which reducing diversification increases both revenue variability and expected revenue (Fig. 4).
Previous work has suggested that changes to fisheries management such as limited entry or
catch share programs may inadvertently act as barrier to diversification and promote 
specializing (Hilborn et al. 2003; Kasperski and Holland 2013). While some of these 
specialized fishing strategies may be associated with increased revenue variability,
the benefit may be that they are also associated with increased revenue.
External market forces may interact in concert with management regulations to promote 
specialization for fishers or other individuals dependent on natural resources. 
In terrestrial agriculture for instance, there are also many instances of specializing 
being profitable; drivers include the ability to specialize on high-profit cash crops
(tobacco, cocoa), the availability of government agricultural subsidies (soybeans, corn in the USA),
and farming on a large industrial scale. With increasing environmental variability and 
extreme events in a changing climate, it is particularly important that individuals who 
rely on natural resources as a primary source of income have strategies to reduce risk.
Though variability in individual revenues may not be actively incorporated into many management
decisions, analyses like ours help to understand the range of diversification strategies 
available and how their benefits may change over time.

@@

##3 ERIC

@@

Whereas a typical hierarchical linear regression model would estimate a single
residual error scale term, $\sigma$, we model the scale of the residual error
with another hierarchical model with a similar form:

>>>>

# Abstract

<!-- TODO slash length of abstract -->

@@

Our results suggest that the benefit of species diversity for revenue
variability is mostly an across-permit effect, not a benefit that can be easily
harnessed by an individual permit holder. Effective methods of diversification
--- purchasing additional permits or switching permits --- are costly and often
limited by regulations such as limited entry programs and individual
transferable quotas (ITQs). Take home ...

@@

Commercial fishers present an interesting example of diversification, given the
extreme risk that some individuals experience, particularly relative to
terrestrial agriculture [@sethi2012b]. The majority of risk insurance programs
available to terrestrial farmers (e.g. crop insurance programs) are not
obtainable by commercial fishers [@kasperski2013]. However, fishers maintain the
ability to lower risk by diversifying their fishing portfolio. They can, for
example, target as many species as permitted by a license, own multiple
licenses, or diversify spatially and fish in multiple regions. The effectiveness
of these diversification strategies for increasing revenue and reducing revenue
variability remains uncertain. For example, theory and empirical studies suggest
there can be substantial benefits to specialization for both farmers and fishers
[@abdulai2000; @arnason2005]. Furthermore, the majority of research examining
the effects of diversification on revenue do so at aggregate levels that combine
individuals (e.g. households, counties, or communities; [@kasperski2013;
@sethi2014]). Thus, key uncertainties remain about the fundamental relationship
between diversity, fishing revenue, and fishing revenue variability at the level
of the individual.

@@

We develop a variance-function regression model [@davidian1987; @hall1989], to
simultaneously models changes in the mean and variance of individual revenue as
a function of covariates. We simultaneously estimate the effect of species
diversity (the inverse of Simpson's diversity index [@simpson1949; @jost2006])
on individual revenue and revenue variability, and estimate effects both within
and across fishing strategies while controlling for fishing effort and latent
time effects (e.g. changing environmental conditions). To our knowledge, this is
the first application of this type of model in any natural resource field.

@@

However, adjusting (increasing or decreasing) an individual's fished species
diversity within the constraints of a permit set tends to be associated with
greater expected variability or uncertainty in year-to-year revenue (Fig.
\ref{fig:within} strategies above 0; Figs. S\ref{fig:fe}, \ref{fig:re},
\ref{fig:mirror}). This effect was strongest for individuals with a halibut
permit or halibut and sablefish permits. Individuals with these strategies
experienced a ~2% increase in variability for a 1% change in the year-to-year
ratio of species diversity (Fig. \ref{fig:within}). Becoming more generalized in
species diversity also leads to increased revenue variability (Fig.
S\ref{fig:mirror}A), suggesting that the act of changing strategies is
a contributor to increased variability of revenue. This benefit of specializing
is most apparent after accounting for fishing effort (Fig. S\ref{fig:noeffort}).

Alaskan halibut and sablefish fisheries were converted to individual fishery
quota (IFQ) management in 1994, which potentially confounds our estimates of the
effect of diversity and a major regulatory change. We refit our models allowing
all strategy-level effects to vary separately before and after the introduction
of IFQs for strategies involving halibut and sablefish permits. The introduction
of IFQs reduced variability by ~25% (XX--XX 95% CI) for individuals with only
halibut or sablefish permits (Fig. S\ref{fig:ifq}A). Furthermore, for most
strategies, the association between changing species diversity and increased
revenue variability became stronger after the introduction of IFQs (Fig.
S\ref{fig:ifq}B). This was especially true for individuals fishing only halibut
or sablefish and less so for individuals combining halibut or sablefish with
a salmon permit (Fig. S\ref{fig:ifq}B).

# Discussion

For individuals deriving livelihoods from natural resources, theory suggests
that adopting strategies with a diversity of assets provide substantial benefits
over strategies with lower diversity [REFS]. However, the ability to effectively
diversify in practice is limited by regulation and market forces. Here we
document the effect of changing species diversity on revenue and revenue
variability across three scale. Each scale reflects individual-level choices
that produce varied levels of species diversity. We find that individuals
fishing permit sets that allow high levels of species diversity tend to have
lower revenue variability. Adding an additional fisheries permit usually
increases revenue and decreases revenue variability. However, within the
constraints of a permit and accounting for effort, specializing (reducing
species diversity) tends to increase revenue and specializing or generalizing
tends to be associated with greater revenue variability, with considerable
differences across fisheries. Thus, the most effective mechanism by which
species diversity can lower revenue variability for individual fishers ---
switching or adding fishing permits --- is often costly, limited by regulations,
and consequently unavailable to many individuals in practice.

<!-- TODO mention the quadrants in figure 4 -->

Our results shows additional support for the idea that individual fishers can
reduce their income variability by diversifying their catch portfolios. However,
most individuals do not appear to change their catch diversity substantially
from year to year, and most strategies adopted by Alaskan fishermen lie in
a high-risk, high-reward quadrant in which reducing diversification increases
both revenue variability and expected revenue. Previous work has suggested that
changes to fisheries management such as limited entry or catch share programs
may inadvertently act as barrier to diversification and promote specializing
(Hilborn et al. 2003; Kasperski and Holland 2013). While some of these
specialized fishing strategies may be associated with increased revenue
variability, the benefit may be that they are also associated with increased
revenue. External market forces may interact in concert with management
regulations to promote specialization for fishers or other individuals dependent
on natural resources. In terrestrial agriculture for instance, there are also
many instances of specializing being profitable; drivers include the ability to
specialize on high-profit cash crops (tobacco, cocoa), the availability of
government agricultural subsidies (soybeans, corn in the USA), and farming on
a large industrial scale. With increasing environmental variability and extreme
events in a changing climate (REF), it is particularly important that
individuals who rely on natural resources as a primary source of income have
strategies to reduce risk. Though variability in individual revenues may not be
actively incorporated into many management decisions, analyses like ours help to
understand the range of diversification strategies available and how their
benefits may change over time.

@@

@@

Whereas traditional hierarchical linear models assume homoscedasticity --- that
the residual variance does not vary systematically --- our variance-function
regression model allows the residual variance to be heteroscedastic and vary as
a function of predictors. Instead of estimating a single residual error scale
term, $\sigma$, we model the scale of the residual error with another
hierarchical model with a similar form:
